\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-05-14 doc/sw-system.tex
Last few edits

However, it is not the case that only these small groups were {\it capable} of
making large contributions. The cumulative distribution of agent skill is shown
in \Fref{fig:crowd:cumulskillplot}: these distributions are significantly
broader than the corresponding distributions of agent contribution, in
\Fref{fig:crowd:cumulplot}. For example,
the most skilled 20\% of agents possess around 80\% of the skill.

The inexperienced volunteers also possess a significant fraction of the skill.
We find that dividing the crowd into ``experienced volunteers'' (who have seen
10 or more training images), and ``inexperienced volunteers'' (the rest),
results in two groups containing approximately equal total skill (see the dashed
curve in \Fref{fig:crowd:cumulskillplot}). The most skillful 20\% of {\it
experienced} volunteers (1824 people) only possess 43\% of the total skill.
The breadth of the skill distributions suggests that  the high level of
contribution made at \SW by experienced volunteers is largely a matter of choice
(or perhaps, availability of time!). There were many other volunteers that  were
skillful enough to make large contributions -- they just didn't classify as many
images.

@@

The similarities between \StageOne and 2, suggest that these numbers could be
scaled to estimate the speed of future \SW projects (or at least those following
a similar blind search strategy). For example, the completion time for a \SW
project may be approximated as
\begin{equation}
    \tau \approx \tau_0 \left(\frac{10^4}{N_p}\right)
                               \left(\frac{N_s}{10^4}\right),
\label{eq:speed}
\end{equation}
where $N_s$ is the number of subjects to be classified, and $N_p$ is the number
of volunteers in the crowd. Just using the numbers above, we find a characteristic
timescale of $\tau_0 = 18\,{\rm days}$, but as we will see in the next section,
we expect this to be significantly shorter in future projects.

@@

What can we learn from the results of the previous section, for future
projects? Potential improvements to the \SW system can be divided into three
categories, performance, efficiency and capacity.

@@

Finally, there could be some performance gains to be made by improving the SWAP
agent model, or its implementation.

Low skill is typically a result  of inexperience (\Sref{sec:results:crowd}) --
but it could be the agent that is inexperienced, as much the classifier.  In a
future paper we plan to investigate the use of the test images as well as the
training images in accelerating the agent's learning (Davis et al, in
preparation). We also plan to investigate the use of offline analysis at
\StageOne, partly for the same reason (see also \Sref{sec:discuss:efficiency}
below.) Given the above findings  about the effects of noise in the SWAP system,
we are also motivated to explore further the possible noise-reducing effects of
the doing the analysis offline (\Sref{sec:results:sample}).

@@

The futuristic lens-finding problem sketched in \Sref{sec:intro}, where $10^5$
lenses are to be found in a $10^4$ square degree wide field imaging survey, is
approximately representative of the challenge facing both the LSST and Euclid
strong lensing science teams \citep{LSSTSciBook,EuclidSciBook}. What role could
citizen scientists at \SW play in lens searching in the next decade? From
\Eref{eq:speed}, and assuming $\tau_0 = 6$ days (as above), we see that for the
\SW crowd to be able to classify $10^8$ images of  photometrically-selected
massive galaxies and groups in approximately 60 days, it would need to contain
$10^7$ volunteers. This is only  a factor of ten larger than the current size of
the Zooniverse userbase.
Alternatively, suppose that automated  lens finding algorithms were able to
select as few as $10^5$ targets for visual inspection, corresponding to a sample
with 10\% purity and a surface density of $\sim10$ per square degree \citep[a
rate within reach of \RF, for example:][]{GavazziEtal2014}. Increasing $\tau_0$
by a factor of three to allow for more inspection time per subject,  we find
that in this case \SW could enable a crowd of just $10^5$ volunteers to assess
them in about three weeks.

@@

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $91-93\%$ and $<1\%$
(Stages I and II). This FPR translates to a purity of approximately 30\%.
We find an even
higher purity at \StageTwo (perhaps as a result of the least visible candidates already
being discarded): here, 100\% purity was reached at just under $90\%$
completeness. Because \SW is a supervised learning system, these numbers should
be taken to be upper limits to what we should expect for real lenses (which have
not been selected for a high visibility training set).

@@

\Eref{eq:app:infogain} suggests a useful information theoretical
definition of the classifier skill  perceived by the agent. At a fixed
value of $p$, we can take the expectation value of the information gain
$\information$ over  the possible classifications that could be made:
\begin{align}
\langle\information\rangle   =& \sum_C \sum_T \pr(T|C) \pr(C) \log_2 \frac{\pr(T|C)}{\pr(T)} \notag \\
         =& - \sum_T \pr(T) \log_2 \pr(T) \notag \\
          & + \sum_C \pr(C) \sum_T \pr(T|C) \log_2 \pr(T|C) \notag \\
         =&         p  \left[ \mathcal{S}(\CM_{LL}) + \mathcal{S}(1-\CM_{LL}) \right] \notag \\
          &     +(1-p) \left[ \mathcal{S}(\CM_{NN}) + \mathcal{S}(1-\CM_{NN}) \right] \notag \\
          & - \mathcal{S}\left[ p    \CM_{LL}       + (1-p)(1-\CM_{NN})     \right] \notag \\
          & - \mathcal{S}\left[ p (1-\CM_{LL})      + (1-p)   \CM_{NN}      \right]
\end{align}
where $\mathcal{S}(x) = x \log_2{x}$. If we choose to
evaluate $\langle\information\rangle$ at $p = 0.5$, the result has some
useful properties. While random classifiers presented with  $p = 0.5$
subjects have $\skill = 0.0$  as expected, perfect classifiers appear to
the agents to have  $\skill = 1.0$. This suggests that  $\skill$, the
amount of information we expect to  gain when a classifier is presented
with a 50-50 subject, is a reasonable quantification of
\emph{normalised skill}. A consequence of this choice is that the
integrated skill (over all agents' histories) should come out to be
approximately
equal to the number of subjects in the survey, when the search is
``complete'' (and all subjects are fully classified). Therefore, a
particular agent's integrated skill is a reasonable
measure of that classifier's
\emph{contribution} to the lens search.

>>>>

However, it is not the case that only these small groups were {\it capable} of
making large contributions. The cumulative distribution of agent skill is shown
in \Fref{fig:crowd:cumulskillplot}: these distributions are significantly
broader than the corresponding distributions of agent contribution, in
\Fref{fig:crowd:cumulplot}. For example, 80\% of the skill is distributed among
20\% of the agents.
The inexperienced volunteers also possess a significant fraction of the skill.
We find that dividing the crowd into ``experienced volunteers'' (who have seen
10 or more training images), and ``inexperienced volunteers'' (the rest),
results in two groups containing approximately equal total skill (see the dashed
curve in \Fref{fig:crowd:cumulskillplot}): the most skillful 20\% of {\it
experienced} volunteers (1824 people) only possess 43\% of the total skill.
The breadth of the skill distributions suggests that  the high level of
contribution made at \SW by experienced volunteers is largely a matter of choice
(or perhaps, availability of time!). There were many other volunteers that  were
skillful enough to make large contributions -- they just didn't classify as many
images.

@@

The similarities between between \StageOne and 2 regarding the results above,
and despite the difference in the task set,
suggest that these numbers can be scaled to estimate cautiously
the speed of future \SW projects. For example, the completion time for a \SW
project may be approximated as
\begin{equation}
    \tau \approx \tau_0 \left(\frac{10^4}{N_p}\right)
                               \left(\frac{N_s}{10^4}\right),
\label{eq:speed}
\end{equation}
where $N_s$ is the number of subjects to be classified, and $N_p$ is the number
of volunteers in the crowd. Just using the numbers above, we find a characteristic
timescale of $\tau_0 = 18\,{\rm days}$, but as we will see in the next section,
we expect this to be significantly shorter in future projects.

@@

What can we learn from the results of the previous section, for future
projects? Potential improvements to the \SW system can be divided into three
categories: performance, efficiency and capacity.

@@

Finally, there could be some performance gains to be made by improving the SWAP
agent model, or its implementation.
Low skill is typically a result  of inexperience (\Sref{sec:results:crowd}) --
but it could be the agent that is inexperienced, as much the classifier.  In a
future paper we plan to investigate the use of the test images as well as the
training images in accelerating the agent's learning (Davis et al, in
preparation). We also plan to investigate the use of offline analysis at
\StageOne, partly for the same reason (see also \Sref{sec:discuss:efficiency}
below.) Given the above findings  about the effects of noise in the SWAP system,
we are also motivated to explore further the possible noise-reducing effects of
the doing the analysis offline (\Sref{sec:results:sample}).

@@

The futuristic lens-finding problem sketched in \Sref{sec:intro}, where $10^5$
lenses are to be found in a $10^4$ square degree wide field imaging survey, is
approximately representative of the challenge facing both the LSST and Euclid
strong lensing science teams \citep{LSSTSciBook,EuclidSciBook}. What role could
citizen scientists at \SW play in lens searching in the next decade? From
\Eref{eq:speed}, and assuming $\tau_0 = 6$ days (as above), we see that for the
\SW crowd to be able to classify $10^8$ images of  photometrically-selected
massive galaxies and groups in approximately 60 days, it would need to contain
$10^7$ volunteers. This is only  a factor of ten larger than the current size of
the Zooniverse userbase.
Alternatively, suppose that automated  lens finding algorithms were able to
select as few as $10^5$ targets for visual inspection, corresponding to a sample
with 10\% purity and a surface density of $\sim10$ per square degree \citep[a
rate within reach of \RF, for example:][]{GavazziEtal2014}. Now
increasing $\tau_0$
by a factor of three to allow for more inspection time per subject,  we find
that in this case \SW could enable a crowd of just $10^5$ volunteers to assess
them in about three weeks.

@@

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $91-93\%$ and $<1\%$
(in both classification stages).
This FPR translates to a purity of approximately 30\%. We find an even
higher purity at \StageTwo (perhaps as a result of the least visible candidates already
being discarded): here, 100\% purity was reached at just under $90\%$
completeness. Because \SW is a supervised learning system, these numbers should
be taken to be upper limits to what we should expect for real lenses (which have
not been selected for a high visibility training set).

@@

\Eref{eq:app:infogain} suggests a useful information theoretical
definition of the classifier skill  perceived by the agent. At a fixed
value of $p$, we can take the expectation value of the information gain
$\information$ over  the possible classifications that could be made:
\begin{align}
\langle\information\rangle   =& \sum_C \sum_T \pr(T|C) \pr(C) \log_2 \frac{\pr(T|C)}{\pr(T)} \notag \\
         =& - \sum_T \pr(T) \log_2 \pr(T) \notag \\
          & + \sum_C \pr(C) \sum_T \pr(T|C) \log_2 \pr(T|C) \notag \\
         =&         p  \left[ \mathcal{S}(\CM_{LL}) + \mathcal{S}(1-\CM_{LL}) \right] \notag \\
          &     +(1-p) \left[ \mathcal{S}(\CM_{NN}) + \mathcal{S}(1-\CM_{NN}) \right] \notag \\
          & - \mathcal{S}\left[ p    \CM_{LL}       + (1-p)(1-\CM_{NN})     \right] \notag \\
          & - \mathcal{S}\left[ p (1-\CM_{LL})      + (1-p)   \CM_{NN}      \right]
\end{align}
where $\mathcal{S}(x) = x \log_2{x}$. If we choose to
evaluate $\langle\information\rangle$ at $p = 0.5$, the result has some
pleasing properties. While random classifiers presented with  $p = 0.5$
subjects have $\skill = 0.0$  as expected, perfect classifiers appear to
the agents to have  $\skill = 1.0$. This suggests that  $\skill$, the
amount of information we expect to  gain when a classifier is presented
with a 50-50 subject, is a reasonable quantification of
\emph{normalised skill}. A consequence of this choice is that the
integrated skill (over all agents' histories) should come out to be
approximately
equal to the number of subjects in the survey, when the search is
``complete'' (and all subjects are fully classified). Therefore, a
particular agent's integrated skill is a reasonable
measure of that classifier's
\emph{contribution} to the lens search.
